The Effect of Events on Discussion in the Australian Federal Parliament (1901--2017)

### 2018-10-21 outputs/docs/events/paper.Rmd
Update paper

@@

We summarise the days where parliamentary discussion was significantly different from the rest prevailing in that week in Table X. **[Oh boy. Not sure what to do here.]**

@@

We identify differences between governments and elections based on **[INSERT HOW]**.

Of the 36 different governments over this period, we find that 12 **[UPDATE]** of them are significantly different to the government that preceded them. However, two of these -- Page and the first Menzies government -- are likely due to the short length of the Page Government, and should be ignored, leaving only ten **[UPDATE]**. The earliest of these is the second Menzies government starting in 1949, and these ten are concentrated in the second half of our sample, with three of them being in the past twenty years. Similarly, of the 45 general elections that have been held we find that 12 **[UPDATE]** of them define periods that were significantly different to their immediate predecessor. 1974, 1980, 1990, 1998, 2004, and 2007 stand out as elections where the government did not change, but the model suggests there was considerable change in the topics discussed in parliament.

The second Menzies Government was associated with a variety of changes compared with the preceding Chifley Government. The Chifley Government had governed through the end of World War II and difficult economic times that followed. It is also the case that there was a large increase in the number of lower house seats at the 1949 election. Many new politicians entered parliament and this changed representation may also have been party to do with the changed discussion. The sixteen year length of the second Menzies Government, and better economic conditions over this time make it understandable why parliamentary discussion would have been different.

There were six election within the second Menzies Government. However we do not find that any of those elections was associated with a significant change in the topics discussed. In this sense it was a period of consistency, especially when compared with later longer-term governments, such as the Howard Government.

The Menzies Government was succeeded by the Holt Government in January 1966. This is an example where there was a change in government without an election, as the next election only happened in November 1966. The Holt Government is significantly different to the Menzies Government.

@@

**[Add some nice graphs for specific changes.]**

@@

Additionally, the way that we identify unusual periods could be improved. e.g. What happens if an event happens while parliament not sitting then it's not clear how whether that is going to be identiifed. And if an event happens in the middle of an sitting period, but it's really important and changes the discussion of the rest of the period then again that might not be different because of the average comparison.

>>>>

key_events_results <- read_csv(here::here("outputs/results-analysis_model/results-outliers.csv"), col_types = cols())

@@

We summarise the days where parliamentary discussion was significantly different from the rest prevailing in that week in Table \@ref(tab:keyeventsresultstable). **[THIS IS A CUT-DOWN LIST FOR DRAFT PURPOSES - REPLACE WITH ONE FROM IMPROVED MODEL.]**

```{r keyeventsresultstable, echo = FALSE, results = 'asis'}

key_events_results %>%
  kable(booktabs = T, caption = "Days that were significantly different to other days in their sitting period") %>%
  kable_styling(font_size = 12)

@@

We identify differences between neighbouring governments and between neighbouring elections based on calculating 95 per cent credible intervals based on posterior samples. When these do not overlap we consider that the model finds a significant difference between either the neighbouring governments or elections. One advantage of our model over using the STM is that we can restrict the effect of an election to be monotonically decreasing. This allows us to identify differences between government and election effects even when there is a one-term government.

Of the 36 different governments over this period, we find that 12 **[UPDATE BASED ON NEW MODEL]** of them are significantly different to the government that preceded them. However, two of these results -- the significance of the Page  Government and the first Menzies Government -- are likely due to the short length of the Page Government and should be ignored, leaving only ten **[UPDATE BASED ON NEW MODEL]**. The earliest of these is the second Menzies government starting in 1949, and these ten are concentrated in the second half of our sample, with three of them being in the past twenty years **[UPDATE BASED ON NEW MODEL]**. Similarly, of the 45 general elections that have been held we find that 12 **[UPDATE BASED ON NEW MODEL]** of them define periods that were significantly different to their immediate predecessor. 1974, 1980, 1990, 1998, 2004, and 2007 stand out as elections where the government did not change, but the model suggests there was considerable change in the topics discussed in parliament **[UPDATE BASED ON NEW MODEL]**.

The second Menzies Government was associated with a variety of changes compared with the preceding Chifley Government. The Chifley Government had governed through the end of World War II and the difficult economic times that followed. There was also a large increase in the number of seats in the House of Representatives at the 1949 election. Many new politicians entered parliament and this changed representation may also have been partly to do with the changed distribution of discussion topics. The sixteen-year length of the second Menzies Government, and better economic conditions over this time make it understandable why parliamentary discussion would have been different.

There were six elections within the second Menzies Government. However we do not find that any of those elections was associated with a significant change in the topics discussed. In this sense it was a period of consistency, especially when compared with shorter-term Whitlam Government, or the longer-term Howard Government.

The Menzies Government was succeeded by the Holt Government in January 1966. This is an example where there was a change in government without an election, as the next election only happened in November 1966. We find that the Holt Government is significantly different to the Menzies Government. In Figures \@ref(fig:mugovmeziesholt) we compare the topics during the final term of the Menzies Government with the topics of the Holt Government.

```{r mugovmeziesholt, cache = TRUE, echo = FALSE, fig.cap = "Government level effects on selected topics", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-finalmenzies-firstholt.pdf")
```

The Whitlam government is interesting as we find a difference in the topics after it was first elected in December 1972, compared with its second election win in May 1974. Figure \@ref(fig:muelectionwhitlam) compares the topics that are significant in the first Whitlam term and then compares them to those in the second Whitlam government.

```{r muelectionwhitlam, cache = TRUE, echo = FALSE, fig.cap = "Government level effects on selected topics", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-whitlamelections.pdf")
```

The Howard Government is interesting because of the significant differences between elections. For instance, each of the election periods is associated with fairly substantial differences compared with the preceding election periods, and all apart from 2001 are actually significantly different at the 95 per cent level. Figure \@ref(fig:muelectionhoward) compares the topics that are significant in the different Howard terms.

```{r muelectionhoward, cache = TRUE, echo = FALSE, fig.cap = "Different ", out.width = '100%'}
knitr::include_graphics("../../figures/mu_election-howard.pdf")
```

To a certain extent the change after the November 2001 election is expected because of the 9/11 terrorist attacks that had only occurred two months earlier, the Bali Bombings that occurred in October 2002, and the dramatic increase in the discussion related to terrorism and conflict over these years. However the change in 1998 and 2004 is more unexpected. Although the Howard Government is the second-longest serving government and commonly thought of as a period of stability because the senior ministers were consistent as well, it might be that it is better to think of the Howard Government as a combination of three or four different periods and that the Howard Government reinvented itself over this period. **[I WONDER IF THERE'S SOME REFERENCE THAT MAKES SIMILAR POINTS TO THIS?]**

@@

The model also provides estimates of topic prevalence by sitting period. This provides nicely illustrates how the topics change over time, as the daily estimates tend to be quite variable, but using periods defined by governments or elections tend not to provide enough variability For instance, examining Topics 12, 17, 22 and 23, which have to do with war and conflict illustrates Australia's involvement in World War I, World War II, the Korean War, the Vietnam War, the first Gulf War, and the war in Afghanistan and the second Gulf War (Figure \@ref(fig:wartopicsgraph)).

```{r wartopicsgraph, cache = TRUE, echo = FALSE, fig.cap = "Model estimates of topic prevalance by sitting period for those related to war and conflict", out.width = '100%'}
knitr::include_graphics("../../figures/war_stack.pdf")
```

One advantage of our analysis model, compared with using the STM approach is that we can create a measure that is equivalent to testing for outliers in a model where the underlying variables were not latent. We estimate sitting-period level-effects (essentially a median for each topic by sitting period). This can be thought of as as equivalent to a 'actual' level, and then the points for we estimate are the 'estimated' points. The difference defines a measure that can be thought of as essentially a residual and we then identify days where the is at least a 99 per cent difference. This approach means that the model generates dates that are interesting without us having to specify interesting dates.

The model identifies 11 September 2001, 12 October 2002.... **[Add more here.]**

@@

Additionally, the way that we identify unusual periods could be improved. e.g. What happens if an event happens while parliament not sitting then it's not clear how whether that is going to be identified. And if an event happens in the middle of an sitting period, but it's really important and changes the discussion of the rest of the period then again that might not be different because of the average comparison.
